Predicting Survival Probability of NASA Aircraft Engines

### 2025-04-11 index.qmd
Updates to index 4/11

---
title: "Predicting Remaining Useful Life of NASA Aircraft Engines"
subtitle: "Using Cox Proportional Hazards Model"
author: "Jayme Reed & Brad Paton (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  html:
    theme: spacelab
    toc: true
    toc-depth: 3
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

@@

CPH can be used to predict the survival probability at a specific time
$S(t|X=x)$, and the hazard ratio for an individual when compared to a
reference individual $\frac{h(t|X=x)}{h(t|X=x_{ref}}$ [@nahhas2025]. If
the resulting probability is greater than or equal to 50%, it is assumed
the event has not occurred. If the resulting probability is less than
50%, it is assumed the event has occurred. It is also possible to
generate a plot of the estimated survival curves for all values of $t$
that are desired.

@@

Though CPH is typically used with health data, it does have applications
in other fields. This paper will study the survivability of a NASA
aircraft engine using data provided by NASA for the study on propagation
modeling [@saxena2008]. In particular, this paper will be examining the
data from the engine make two which has a training dataset and a testing
dataset.

@@

Below are tables showing the structure of the training and testing data
after the transformations described above.

@@

There are 260 engines in the training data. The histogram shows that the
survival times of the engines are normally distributed and the box plot
shows there six outliers in the dataset. The engines have a median
survival time of 199 cycles with a standard deviation of 46.8. The
maximum number of cycles an engine lasted was 378 while the minimum
number of cycles an engine lasted was 128.

```{r, echo = FALSE}
train_grouped <- train %>% 
  group_by(id) %>%
  slice_max(time)

@@

colnames(summarymetrics) <- c("Metric", "Value")
summarymetrics$Value <- round(summarymetrics$Value, digits = 2)
kable(summarymetrics, format = "markdown")

@@

To determine the initial survival function, the Kaplan-Meier survival
curve was generated and is shown below. The graph visualizes the
survival curve and the information that was gathered through the
histogram.

@@

**Covariate Selection**

Based on the information provided by NASA on their data, there are
various sensors and operational setting fields that do not impact the
survival rate of an engine. To determine which covariates will impact
the survival probability, the below univariate test was run. The
covariates in which the p-value is $< 0.05$ will be considered for the
CPH models while the other covariates will be excluded.

@@

Based on the above test, the covariates that will be considered for the
CPH models are os1, os3, sm1, sm5, sm6, sm7, sm11, sm12, sm13, sm14,
sm19, sm20, and sm21 which is fourteen out of twenty-four possible
covariates.

Generation of all CPH models will be done using R, specifically the R
package `survival`. To generate the CPH model with properly formatted
survival data, a survival object is created by running
`Surv(time, event)` where `time` is the time field and `event` is the
event field. To generate the actual model, the survival object is then
applied to the function `coxph(survival object ~ x, data)` where `x`
represents one or multiple covariate(s) and `data` is the survival data
that is being modeled.[@R-base].

The table below provides the model number, the covariates used, and the
concordance index. The concordance index generates a value to determine
which model is the most accurate with the provided covariates. After
determining the most accurate model, the CPH assumptions will be tested.

@@

cox1 <- coxph(data = data_grouped, surv_object ~ os1 + os3 + sm1 + sm5 + sm6 + sm7 + sm11 + sm12 + sm13 + sm14 + sm19 + sm20 + sm21)

cox2 <- coxph(data = data_grouped, surv_object ~ os1 + os3 + sm1 + sm5)

cox3 <- coxph(data = data_grouped, surv_object ~ os1 + os3 + sm1 + sm5 + sm6 + sm7 + sm13 + sm14 + sm19 + sm21)

cox4 <- coxph(data = data_grouped, surv_object ~ sm5 + sm6 + sm7 + sm11 + sm12 + sm13 + sm14 + sm19 + sm20 + sm21)

cox5 <- coxph(data = data_grouped, surv_object ~ os1 + os3 + sm1 + sm5  + sm7 + sm11 + sm12 + sm13 + sm19 + sm20 )

cox6 <- coxph(data = data_grouped, surv_object ~ os1  + sm1 + sm5  + sm7 + sm11 + sm12 + sm13 + sm19  + sm21)

cox7 <- coxph(data = data_grouped, surv_object ~ os1  + sm13 + sm14 + sm19 + sm20)

cox8 <- coxph(data = data_grouped, surv_object ~ os1 + os3 + sm21)

cox9 <- coxph(data = data_grouped, surv_object ~ os1 + os3 + sm5 + sm6 + sm12 + sm13  + sm19 + sm20 + sm21)

cox10 <- coxph(data = data_grouped, surv_object ~  os3 + sm1 + sm5 + sm6  + sm11 + sm12 + sm13  + sm19 + sm20 + sm21)

@@

```{r}
concordance_results <- data.frame(c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5", 
                                    "Model 6", "Model 7", "Model 8", "Model 9", "Model 10"), 
                   c("os1, os3, sm1, sm2, sm5, sm6, sm7, sm11, sm12, sm13, sm14, sm19, sm20, sm21", 
                     "os1, os3, sm1, sm2, sm5", 
                     "os1, os3, sm1, sm5, sm6, sm7, sm13, sm14, sm19, sm21", 
                     "sm5, sm6, sm7, sm11, sm12, sm13, sm14, sm19, sm20, sm21", 
                     "os1, os3, sm1, sm2, sm5, sm7, sm11, sm12, sm13, sm19, sm20", 
                     "os1, sm1, sm2, sm5, sm7, sm11, sm12, sm13, sm19, sm21",
                     "os1, sm13, sm14, sm19, sm20", 
                     "os1, os3, sm21",
                     "os1, os3, sm2, sm5, sm6, sm12, sm13, sm19, sm20, sm21",
                     "os3, sm1, sm2, sm5, sm6, sm11, sm12, sm13, sm19, sm20, sm21"),
                   c(cox1$concordance[6], cox2$concordance[6], cox3$concordance[6], cox4$concordance[6], cox5$concordance[6], cox6$concordance[6], cox7$concordance[6], cox8$concordance[6], cox9$concordance[6], cox10$concordance[6]))

colnames(concordance_results) <- c("Model", "Covariates", "Concordance Index")

kable(concordance_results, format = "markdown")

@@

Model 1 had the highest concordance index and used all of the covariates
that were determined to be significant to the survival of the engine.
The proceeding analysis will be done using model 1. The below tables
provide the summary values from the model.

```{r}
kable(tidy(cox1), format = "markdown")
g <- glance(cox1)
g <- g %>% select(n, nevent, concordance, std.error.concordance, logLik, statistic.log, p.value.log, statistic.wald, p.value.wald, statistic.sc, p.value.sc)
kable(g, format = "markdown")
```

@@

```{r}
plot_model(cox1,
           dot.size = 2,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratios")+
  ylab(label = "Covariates")+
  xlab(label = "Estimated Effect on Hazard Ratio")
```

@@

The plots of the Martingale residuals vs the range of values for each
covariate is shown below. For each covariate, there is no clear slop
indicating the linearity assumption is true for the covariates in the
model.

```{r}
par(mfrow = c(4,4),
      mar = c(2,2,4,1))

martingale <- resid(cox1, type = "martingale")

@@

@@

@@

@@

@@

The `cox.zph` function can be used to check the proportional hazards
assumption. As mentioned above, the function calculates p values for
each covariate. The global p-value will first be compared to
$\alpha = 0.05$. If that p-value is $<\alpha$, then at least one
covariate violates the PH assumption and the model will need to be
adapted.

The output below shows the p-values for the hypothesis test and
Schoenfeld residuals. The global p-value is $< \alpha$ so at least one
covariate violates the PH assumption. Examining each individual
covariate through both p-values and the plot of the Schoenfeld residuals
shows that the covariates os1, os3, sm1, sm5, sm13, and sm14 all violate
the PH assumption.

```{r}
cph_ph <- cox.zph(cox1) %>%
  pluck("table") %>%
  as.data.frame() %>%
  tibble::rownames_to_column("variable") %>%
  as_tibble()
kable(cph_ph, format = "markdown")
```

```{r}
# Function to put ph_plots on one visual
create_ph_plots <- function(coxmodel, row, col){
  # Function takes argument that is a cox model from coxph()

@@

create_ph_plots(cox1, 3, 5)
```

@@

The results of the above assumptions demonstrated that several
covariates violate the PH assumption. As such, the model will be
recreated by removing the covariates that violate the PH assumption.
After recreating the model, the PH assumption and the linearity
assumption will be tested again.

The table below demonstrates the new model information which has seven
covariates: sm2, sm5, sm7, sm11, sm12, sm20, and sm21.

```{r}
cox1.1 <- coxph(data = data_grouped,Surv(time, status) ~ sm2 + sm5 + sm7 + sm11 + sm12 + sm20 + sm21)

kable(tidy(cox1.1), format = "markdown")
g <- glance(cox1.1)
g <- g %>% select(n, nevent, concordance, std.error.concordance, logLik, statistic.log, p.value.log, statistic.wald, p.value.wald, statistic.sc, p.value.sc)
kable(g, format = "markdown")

@@

The output below shows the p-values for the new model's PH assumption
hypothesis test and Schoenfeld residuals. The global p-value is
$> \alpha$ so the PH assumption has been met for the model even though
the covariate sm5 has an individual p-value of $< \alpha$. As such, this
new model does not violate the PH assumption and the analysis can
continue.

```{r}
ph_2 <- cox.zph(cox1.1) %>%
  pluck("table") %>%
  as.data.frame() %>%
  tibble::rownames_to_column("variable") %>%
  as_tibble()

@@

create_ph_plots(cox1.1, 3, 5)

@@

Since a new model was generated, the below plots confirm that the
linearity assumption is still met for this model.

```{r}

par(mfrow = c(3,3),
      mar = c(2,2,4,1))

martingale <- resid(cox1.1, type = "martingale")

plot(data_grouped$sm21, martingale, main = "Martingale Residuals: sm21")
abline(h = 0, col = "red")

### Model Applications and Predictions

The model created can be used to calculate several relevant metrics when
predicting remaining time for the engines until expected failure.
Predictions will be made on the largest value for time for each
individual engine.

#### Model Coefficients

@@

#### Hazard Rate Calculation

The coefficients above can be used to calculate the hazard rate at time
$t$. The hazard rate will be calculated for each engine at the engines
max time value.

The engines with the higher hazard rates are at a higher risk of
experiencing failure.

```{r}
data_grouped_reduced <- data_grouped %>% select(id, time, sm2, sm5, sm7, sm11, sm12, sm20, sm21, status)

@@

#### Cumulative Hazard Calculation

The cumulative hazard represents the total hazard accumulated for the
individual over time and like the hazard rate, is another measure that
can be used to estimate the risk for the individual at a point in time.
The cumulative hazard for each individual is shown below. Higher values
represent a higher chance of failure.

$H(t) = \int_0^t h(t) dt$ where, $H(t)$ is the cumulative hazard from
time 0 to time $(t)$ $h(t)$ is the hazard rate at time $t$

The cumulative hazard of individuals in the model are shown in the visual below. Higher cumulative hazard indicates a higher level of risk at time `t`.

```{r}
sf <- survfit(cox1.1)

@@

#### Survival probability

While cumulative hazard can be used by itself to give an idea of an
individual's risk, it can also be used to calculate a probability of
survival.

Probability Calculation: $S(t) = e^{-H(t)}$ $H(t)$ is the cumulative
hazard up to time $t$.

```{r}
data_grouped_reduced$SurvivalProbability <- predict(cox1.1, type = "survival", newdata = data_grouped_reduced)
kable(head(data_grouped_reduced))

The resulting survival curve from the model is shown below.

```{r}

survtimes <- data.frame(Percent = c("100% Survival", "< 25% Survival,", "< 50% Survival", "< 75% Survival", "<90% Survival"),
                        Time = c(sf$time[sf$lower < 1][1], 
                                    sf$time[sf$lower < .75][1], 
                                    sf$time[sf$lower < .5][1], 
                                    sf$time[sf$lower < .25][1], 
                                    sf$time[sf$lower < .1][1]))

@@

The implications of an engine failure are severe, with a possible result including an aircraft crash with a loss of human life and millions of dollars. Based on these implications, there is no acceptable room for failure for aircraft engines. The model created a confidence interval for estimating probabilities at each value of time. The values for time used will be the value where the probability of survival at the lower end of the confidence interval to exercise caution.

```{r}
kable(survtimes)

@@

The lower end of the confidence interval for survival probability is shown to hold stable at 100% until `t = 128` with a gradual decrease afterwards. Half of the engines are expected to survive until `t = 221` and 90% of the engines are expected to fail at `t = 316`. These times are reflected by the survival curve shown above. The cumulative hazard over time plot confirms the hazard over time increases. The plot shows a low hazard until `t = 128` with a gradual increase.

>>>>

---
title: "Predicting Survival Probability of NASA Aircraft Engines"
subtitle: "Using Cox Proportional Hazards Model"
author: "Jayme Reed & Brad Paton (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  html:
    theme: spacelab
    toc: true
    toc-depth: 3
    code-fold: true
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

@@

CPH can be used to predict the survival probability at a specific time
$S(t|X=x)$, and the hazard ratio for an individual when compared to a
reference individual $\frac{h(t|X=x)}{h(t|X=x_{ref}}$ [@nahhas2025]. The
equation is

$$
S(t) = e^{-H(t)}
$$ {#eq-survprob}

where, $H(t)$ is the cumulative hazard up to time $t$.

If the resulting probability is greater than or equal to 50%, it is
assumed the event has not occurred. If the resulting probability is less
than 50%, it is assumed the event has occurred. It is also possible to
generate a plot of the estimated survival curves for all values of $t$
that are desired. However, CPH is not able to predict the specific time
the event is going to occur which is considered a limitation of the
model.

@@

Though CPH is typically used with health data, it does have applications
in other fields. This paper will study the survival probability of a
NASA aircraft engine using data provided by NASA for the study on
propagation modeling [@saxena2008]. In particular, this paper will be
examining the data from the engine design two which has a training
dataset and a testing dataset.

@@

The table below shows the structure for the combined training and
testing data after the transformations described above.

@@

There are 519 engines in the combined data with 260 engines in the
training data and 259 engines in the testing data. The below metrics are
generated using the training data as that data contains no censored
data. The histogram shows that the survival times of the engines are
normally distributed and the box plot shows there six outliers in the
dataset. The engines have a median survival time of 199 cycles with a
standard deviation of 46.78. The maximum number of cycles an engine
lasted was 378 while the minimum number of cycles an engine lasted was
128. Since the testing data contains all censored data, the basic
metrics were not generated.

```{r}
train_grouped <- train %>% 
  group_by(id) %>%
  slice_max(time)

@@

colnames(summarymetrics) <- c("Metric", "Value")
summarymetrics$Value <- round(summarymetrics$Value, digits = 2)
kable(summarymetrics, format = "markdown", align = "c")

@@

To determine the initial survival function, using the training data, the
Kaplan-Meier survival curve was generated and is shown below. The graph
visualizes the survival curve and the information that was gathered
through the histogram.

@@

Generation of the CPH models will be done using R, specifically the R
package `survival`. To generate the CPH model with properly formatted
survival data, a survival object is created by running
`Surv(time, event)` where `time` is the time field and `event` is the
event field. To generate the actual model, the survival object is then
applied to the function `coxph(survival object ~ x, data)` where `x`
represents the covariate(s) and `data` is the survival data that is
being modeled [@R-base].

Based on the information provided by NASA on their data, there are
various sensors and operational setting fields that do not impact the
survival rate of an engine. To determine which covariates are
statistically significant, the below univariate test was run. The test
uses the Wald test to determine which covariates will significantly
impact the model individually. The covariates in which the p-value is
$< 0.05$ will be considered statistically significant.

@@

Based on the above test, none of the covariates are statistically
significant individually with all the data. As such, to determine the
best model to use for this analysis, a forward and backward step-wise
regression will be done to generate the models with the best Akaike
Information Criterion (AIC) and Bayesian Information Criterion (BIC)
values and compare the concordance values of those models with models
that have covariates that were statistically significant when used in a
CPH model with the training data. The model with the highest concordance
value, and therefore, the most accurate will be used for the analysis.

The table below provides the model number, the covariates used, the AIC
and BIC from the step-wise regression if applicable, and the concordance
index. The concordance index generates a value to determine which model
is the most accurate with the provided covariates. After determining the
most accurate model, the CPH assumptions will be tested.

@@

all_cov <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm16 + sm17 + sm18 + sm19 + sm20 + sm21)

aic <- MASS::stepAIC(all_cov, trace = 0)
n <- nrow(data_grouped)
bic <- step(all_cov, direction = "both", k = log(n), trace = 0)

train_sig <- lapply( univ_formulas, function(x){coxph(x, data = train_grouped)})

cox1 <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm16 + sm17 + sm18 + sm19 + sm20 + sm21)

cox2 <- coxph(data = data_grouped, surv_object ~ os3 + sm3 + sm4 + sm5 + sm8 + sm9 + sm15 + sm16 + sm17 + sm18 + sm20)

cox3 <- coxph(data = data_grouped, surv_object ~ os3 + sm3 + sm4 + sm8 + sm9 + sm15 + sm18)

cox4 <- coxph(data = data_grouped, surv_object ~ os3 + sm13 + sm14 + sm19)

@@

```{r, echo = FALSE}
stepwise_results <- data.frame(c("model 1", "model 2", "model 3", "model 4"), 
                               c("All covariates", "os3, sm3 - sm5, sm8 - sm9, sm15 - sm18, sm20", "os3, sm3 - sm4, sm8-sm9, sm15, sm18", "os3, sm13 - sm14, sm19"),
                               c("2449.92", "2431.88", "N/A", "N/A"),
                               c("2547.71", "N/A", "2461.31", "N/A"),
                               c(cox1$concordance[6], cox2$concordance[6], cox3$concordance[6], cox4$concordance[6]))

colnames(stepwise_results) <- c("Model", "Covariates", "AIC", "BIC", "Concordance")

kable(stepwise_results, format = "markdown")

@@

Based on the above model results, the model that has the highest
concordance value is model 1 which has all of the covariates. However,
before proceeding with that model, a test for linear dependency will be
run to determine all covariates are linearly independent.

```{r}
kable(alias(cox1), format = "markdown")
```

The above result indicates that sm19 needs to be removed from the model
as it linearly dependent on another value. In addition, sm16 resulted in
a possible infinite coefficient so that covariate has also been removed.
The proceeding analysis will be done using model 1 without sm16 and
sm19. The below tables provide the summary values from the model.

```{r}
cox1.1 <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm17 + sm18 + sm20 + sm21)

ci <- cox1.1$concordance[6]
```

```{r, echo = FALSE}
kable(tidy(cox1.1), format = "markdown")

g <- glance(cox1.1)
g <- g %>% select(n, nevent, concordance, std.error.concordance, logLik, statistic.log, p.value.log, statistic.wald, p.value.wald, statistic.sc, p.value.sc)

kable(g, format = "markdown")
```

@@

```{r, echo = FALSE}
plot_model(cox1.1,
           dot.size = 2,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratios")+
  ylab(label = "Covariates")+
  xlab(label = "Estimated Effect on Hazard Ratio")
```

@@

The plots of the Martingale residuals vs the range of values for each
covariate is shown below. For each covariate, there is no clear slope
indicating the linearity assumption is true for the covariates in the
model.

```{r}
martingale <- resid(cox1.1, type = "martingale")
```

```{r, echo = FALSE}
par(mfrow = c(4,4),
      mar = c(2,2,4,1))

@@

plot(data_grouped$os3, martingale, main = "Martingale Residuals: os3")
abline(h = 0, col = "red")

@@

plot(data_grouped$sm3, martingale, main = "Martingale Residuals: sm3")
abline(h = 0, col = "red")

plot(data_grouped$sm4, martingale, main = "Martingale Residuals: sm4")
abline(h = 0, col = "red")

@@

plot(data_grouped$sm8, martingale, main = "Martingale Residuals: sm8")
abline(h = 0, col = "red")

plot(data_grouped$sm9, martingale, main = "Martingale Residuals: sm9")
abline(h = 0, col = "red")

plot(data_grouped$sm10, martingale, main = "Martingale Residuals: sm10")
abline(h = 0, col = "red")

@@

plot(data_grouped$sm15, martingale, main = "Martingale Residuals: sm15")
abline(h = 0, col = "red")

plot(data_grouped$sm17, martingale, main = "Martingale Residuals: sm17")
abline(h = 0, col = "red")

plot(data_grouped$sm18, martingale, main = "Martingale Residuals: sm18")
abline(h = 0, col = "red")

@@

The `cox.zph` function can be used to check the proportional hazards
assumption. As mentioned, the function calculates p values for each
covariate. The global p-value will first be compared to $\alpha = 0.05$.
If that p-value is $<\alpha$, then at least one covariate violates the
PH assumption and the model will need to be adapted.

The output below shows the p-values for the hypothesis test and
Schoenfeld residuals. The global p-value is $> \alpha$ so none of the
covariates violates the PH assumption on a global level. Examining each
individual covariate through both p-values and the plot of the
Schoenfeld residuals does show that os1, os3, sm13, and sm14 violate the
PH assumption. To ensure the model is as accurate as possible, the model
will addapted without those covariates and the global p-values will be
compared.

```{r}
cph_ph <- cox.zph(cox1.1) %>%
  pluck("table") %>%
  as.data.frame() %>%
  tibble::rownames_to_column("variable") %>%
  as_tibble()
kable(cph_ph, format = "markdown")
```

```{r, echo = FALSE}
# Function to put ph_plots on one visual
create_ph_plots <- function(coxmodel, row, col){
  # Function takes argument that is a cox model from coxph()

@@

create_ph_plots(cox1.1, 3, 5)
```

@@

The results of the above assumptions demonstrated that several
covariates violate the PH assumption on a local level. As such, the
model will be recreated by removing the covariates that violate the PH
assumption. After recreating the model, the PH assumption and the
linearity assumption will be tested again with the new PH global p-value
being compared to the original model's global p-value.

The table below demonstrates the new model information which has
eighteen covariates: os2, sm1 - sm12, sm15, sm17, sm18, sm20, sm21.

```{r}
cox1.2 <- coxph(data = data_grouped, surv_object ~ os2 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm15 + sm17 + sm18 + sm20 + sm21)

kable(tidy(cox1.2), format = "markdown")
g <- glance(cox1.2)
g <- g %>% select(n, nevent, concordance, std.error.concordance, logLik, statistic.log, p.value.log, statistic.wald, p.value.wald, statistic.sc, p.value.sc)
kable(g, format = "markdown")

@@

The output below shows the p-values for the new model's PH assumption
hypothesis test and Schoenfeld residuals. The global p-value is now
$< \alpha$ so the PH assumption has not been met for the new model. As
such,, the remaining analysis will be done using the first model that
had a global p-value $>\alpha$ which has all but two covariates.

```{r, echo = FALSE}
ph_2 <- cox.zph(cox1.2) %>%
  pluck("table") %>%
  as.data.frame() %>%
  tibble::rownames_to_column("variable") %>%
  as_tibble()

@@

create_ph_plots(cox1.2, 3, 5)

@@

If the new model had passed the PH assumption, the linearity assumption
would have been checked again.

### Model Results

**Evaluation**

The CPH model that was selected has a concordance value of `{r} ci`
which indicates that it is more than 50% accurate in predicting the
survival of the machines. However, this model is not completely accurate
so there is an expectation of some error in the model survival
probabilities which will lead to caution in providing results in
application settings.

**Calculations**

@@

The coefficients above can be used to calculate the hazard rate
[@eq-hazardrate], the cumulative hazard value[@eq-cumulativehr], and the
survival probability [@eq-survprob]. The hazard rate at time $t$ will be
calculated for each engine at the engines max time value. The cumulative
hazard value is the total hazard accumulated for the subject over time
and is used to estimate the risk for the subject at a point in time. The
engines that have higher hazard rates or higher cumulative hazard values
are at a greater risk of experiencing failure. The below table provides
the hazard rate and suvivial probability for the first 10 engines in the
dataset.

```{r}
data_grouped_reduced <- data_grouped %>% select(id, time, os1, os2, os3, sm1, sm2, sm3, sm4, sm5, sm6, sm7, sm8, sm9, sm10, sm11, sm12, sm13, sm14, sm15, sm17, sm18, sm20, sm21, status)

@@

data_grouped_reduced$SurvivalProbability <- predict(cox1.1, type = "survival", newdata = data_grouped_reduced)

print_data <- data_grouped_reduced %>% select(id, time, status, HazardRate, SurvivalProbability)

kable(head(print_data, n = 10), format = "markdown")
```

The below graphs demonstrate the cumulative hazard over time and the
survival curve from the prediction model.

```{r, echo = FALSE}
sf <- survfit(cox1.1)

@@

```{r, echo = FALSE}
survtimes <- data.frame(Percent = c("100% Survival", "75% Survival,", "50% Survival", "25% Survival", "10% Survival"),
                        Time = c(sf$time[sf$lower < 1][1], 
                                    sf$time[sf$lower < .75][1], 
                                    sf$time[sf$lower < .5][1], 
                                    sf$time[sf$lower < .25][1], 
                                    sf$time[sf$lower < .1][1]))

@@

The implications of an engine failure are severe, as failure can lead to
injury or death and the loss of millions of dollars. As such,
understanding when an engine will fail allows for maintenance to occur
before negative impacts occur. While CPH does not allow for the
prediction of specific failure times, it does allow for the creation
confidence intervals which can be used to gain an understanding of when
engines are most likely to fail. The table below displays the percent of
survival at the lower end of the confidence interval to exercise
caution.

```{r}
kable(survtimes, format = "markdown")

@@

The lower end of the confidence interval for survival probability is
shown to hold stable at 100% until `t = 128` with a gradual decrease
afterwards. Half of the engines are expected to survive until `t = 221`
and 90% of the engines are expected to fail at `t = 316`. These results
can be viewed in both the above survival graph and the cumulative hazard
plot which shows a low hazard until `t = 128` with a gradual increase.
